---
title: "WIP: Suffolk Annual Waste Monitoring Report 21/22"
author: 
    - Deborah Sacks, Sacks Consulting
    - Thomas Zwagerman, Suffolk County Council
scctemplate:
  header:
    site_branding: "Suffolk County Council"
  navigation:
    breadcrumb_trail:
      - href: "index.html"
        text: "WIP: SWAMR"
editor_options: 
  markdown: 
    wrap: sentence
---

\newpage

# Introduction

The Suffolk Minerals & Waste Local Plan was adopted in July 2020.
The Plan contains a Vision explaining how the County will meet its statutory obligation for the supply of aggregates and the sustainable management of waste.

It contains policies for determining planning application for minerals and waste development and identifies sites for future development for these uses.

It is important to understand whether the policies in the Plan are being delivered as intended and the extent to which the waste arisings and movements that were forecast are correct.
Because of the nature of waste data, such forecasts will never be exact, but trends can be identified and broad conclusions reached.

The Plan identifies sites that are suitable for the management of waste and this Monitoring Report will examine whether such sites have been taken forward for this use, or whether the policies should be amended in a future Plan.

The Local Plan contains a Policy Monitoring Framework which identifies the Performance Indicators that will be used to monitor the Plan which are the number of times that the relevant policies are triggered in the decision-making process.

Suffolk County Council also has a Development Management Local Monitoring and Enforcement Plan, adopted in 2016.
This sets out how the Council deals with the monitoring of developments as they are delivered including site visits to ensure that planning conditions are being carried out as intended.

This report looks at the effectiveness of the waste policies in delivering the policy outcomes in the Plan.

In particular it examines whether Suffolk is net self-sufficient in the management of the waste arising with the Plan Area or whether additional waste management facilities are required to achieve this.

## Coronavirus (COVID-19): The impact of the pandemic on local authority waste collection and services

When reading this report, and especially when comparing annual totals between 2020/21 and 2021/22, the context of the pandemic should be taken into account.

The COVID-19 pandemic, which commenced in March 2020, caused disruption to local authority waste collection services. Services were heavily disrupted during April to June 2020.

As restrictions eased through the summer and into the autumn of 2020, local authorities and businesses acclimatised to working under lockdown conditions, and there was less disruption to services in the remainder of 2020.

During 2021 lockdown restrictions were eased and despite further waves of infection there was less disruption to waste services.

Wider issues, such as the state of the economy during the pandemic will also have had an effect on the amount of waste generated.

A [full timeline of the UK government coronavirus lockdown](https://www.gov.uk/government/statistics/local-authority-collected-waste-management-annual-results-202122/local-authority-collected-waste-management-annual-results-202122#coronavirus-covid-19-the-impact-of-the-pandemic-on-local-authority-waste-collection-and-services) is available. A more in-depth analysis of the impact of the COVID-19 pandemic on waste services is given in the [Local authority collected waste management - annual results 2021/22 report](https://www.gov.uk/government/statistics/local-authority-collected-waste-management-annual-results-202122/local-authority-collected-waste-management-annual-results-202122#coronavirus-covid-19-the-impact-of-the-pandemic-on-local-authority-waste-collection-and-services) from the Department for Environment Food & Rural Affairs (DEFRA).


# Methodology

## Data Sources
The [Waste Data Interrogator 2019](https://find-data-beta.cloudapps.digital/dataset/d409b2ba-796c-4436-82c7-eb1831a9ef25/2019-waste-data-interrogator), [2020](https://www.data.gov.uk/dataset/bb40d091-a346-4b75-aa54-df7d347bed93/2020-waste-data-interrogator) and [2021](https://www.data.gov.uk/dataset/d8a12b93-03ef-4fbf-9a43-1ca7a054479c/2021-waste-data-interrogator) that have been produced by the Environment Agency is the primary data source for the report.

The second source is Local Authority Collected Waste data Defra publish as part of the Waste Management Statistics and information from the Waste Data Flow system, which can be found under [ENV18 - Local Authority Collected Waste](https://www.gov.uk/government/statistical-data-sets/env18-local-authority-collected-waste-annual-results-tables-202122). These annual results are published alongside a [national statistical report](https://www.gov.uk/government/statistics/local-authority-collected-waste-management-annual-results-202122/local-authority-collected-waste-management-annual-results-202122#coronavirus-covid-19-the-impact-of-the-pandemic-on-local-authority-waste-collection-and-services).

Finally, [Remaining Landfill Capacity 2019-2021](https://www.data.gov.uk/dataset/237825cb-dc10-4c53-8446-1bcd35614c12/remaining-landfill-capacity) reported by the Environment Agency is also used.

## Code

This report was generated in R Markdown. The source code for this report, all figures and calculations are provided in a [Github repository](https://github.com/SCC-Planning/wamreport). Full documentation is provided, and will allow other local authorities to generate their own annual waste monitoring report by adapting the code.

```{r setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE) 
options(scipen=999)
```

```{r reading_wdi, include = FALSE}
lacw <- wamreport$lacw
# Reading in and combining relevant data
districts <- c(
  "Ipswich Borough Council",
  "Mid Suffolk District Council",
  "East Suffolk Council",
  # Waveney and SCDC merged into East in 19-20
  "Waveney District Council",
  "Suffolk Coastal District Council",
  # Forest Heath and St Edmunsbury mered into West in 19-20
  "West Suffolk Council",
  "Forest Heath District Council",
  "St Edmundsbury Borough Council"
)

lacw_suffolk <- lacw |>
  dplyr::filter(`Local Authority` %in% districts)

# Changing names to allow aggregating of former District Councils
lacw_suffolk$`Local Authority` <- gsub("Waveney District Council", "East Suffolk Council", lacw_suffolk$`Local Authority`)
lacw_suffolk$`Local Authority` <- gsub("Suffolk Coastal District Council", "East Suffolk Council", lacw_suffolk$`Local Authority`)

lacw_suffolk$`Local Authority` <- gsub("Forest Heath District Council", "West Suffolk Council", lacw_suffolk$`Local Authority`)
lacw_suffolk$`Local Authority` <- gsub("St Edmundsbury Borough Council", "West Suffolk Council", lacw_suffolk$`Local Authority`)

lacw_suffolk <- lacw_suffolk |>
  dplyr::select(-Region, -`ONS Code`, -`JPP order`, -`Authority type`) |>
  dplyr::group_by(`Financial Year`, `Local Authority`) |>
  dplyr::summarise(across(everything(), sum), .groups = "drop")

lacw_scc <- lacw |>
  dplyr::filter(`Local Authority` == "Suffolk County Council")

lacw_scc_21 <- lacw_scc |>
  dplyr::filter(`Financial Year` == "2021-22") |>
  dplyr::select(`Local Authority`,`Total local authority collected waste (tonnes)`)

lacw_suffolk_21 <- lacw_suffolk |>
  dplyr::filter(`Financial Year` == "2021-22") |>
  dplyr::select(`Local Authority`,`Total local authority collected waste (tonnes)`)
```

# Waste Managed in Suffolk

## Local Authority Collected Waste

The total amount of Local Authority Collected Waste (LACW) in Suffolk, after a reduction dip in 2020-21, has increased again to a similar level as before the pandemic as shown in Figure \@ref(fig:totallocalauthority):

```{r totallocalauthority,  fig.alig = "center", fig.height = 4, fig.width = 8, fig.cap = "\\label{fig:totallocalauthority}Source: [ENV18 - Local Authority Collected Waste](https://www.gov.uk/government/statistical-data-sets/env18-local-authority-collected-waste-annual-results-tables)"}

p1 <- sccthemes::scc_barchart(
  lacw_scc,
  "Financial Year",
  "Total local authority collected waste (tonnes)",
  title = "Total local authority collected waste (tonnes)"
)

p1
```

Household waste arising has also returned to a level similar to that before 2020-21, as shown in Figure \@ref(fig:householdwaste)

This is in line of expectation, and similar to the national picture, which saw household waste increase with the easing of lockdown in Spring 2021.

Consumption overall may have reduced as a result of lower economic activity in 2020-21, and returned to previous levels as economic activity has ramped up again in 2021-22.

```{r householdwaste, fig.alig = "center", fig.height = 4, fig.width = 8, fig.cap = "Source: [ENV18 - Local Authority Collected Waste](https://www.gov.uk/government/statistical-data-sets/env18-local-authority-collected-waste-annual-results-tables)"}
p2 <- sccthemes::scc_barchart(
  lacw_scc,
  "Financial Year",
  "Household - total waste (tonnes)",
  title = "Household - total waste (tonnes)"
)

p2
```

## Recycling Rates

The amount of LACW sent for recycling has steadily reduced between 2015-16 and 2020-21, as shown in Figure \@ref(fig:sentforrecycling).
In 2021-22 however, there is again an increase in the amount of household waste set for recycling/composting/reuse.
It should be noted however that it is still a smaller when compared to 2019-20, pre-pandemic.

```{r sentforrecycling, fig.alig = "center", fig.height = 4, fig.width = 8, fig.cap = "Source: [ENV18 - Local Authority Collected Waste](https://www.gov.uk/government/statistical-data-sets/env18-local-authority-collected-waste-annual-results-tables)"}
p3 <- sccthemes::scc_barchart(
  lacw_scc,
  "Financial Year",
  "Household - waste sent for recycling-composting-reuse (tonnes)",
  title = "Household - waste sent for recycling-composting-reuse (tonnes)"
)

p3
```

Suffolk historically had one of the highest recycling rates in England but this has reduced in all the Suffolk Districts over the past seven years as shown in Table \@ref(tab:recyclingrates). It should be noted that the recycling rate in Suffolk is now lower than the official recycling rate in England as a whole (44.1%).

Compared to last year, the recycling rate as a percentage has stayed constant.

This indicates that, although in absolute tonnage more household waste has been sent for recycling/composting/reuse in 2021-22 \@ref(fig:sentforrecycling), this is a result the total local authority collected waste increasing and not due to an increase in the recycling rates.

```{r recyclingrates, ft.align = "center", tab.id = "recyclingrates", label = "recyclingrates", tab.cap = "Source: [ENV18 - Local Authority Collected Waste](https://www.gov.uk/government/statistical-data-sets/env18-local-authority-collected-waste-annual-results-tables)"}

recycling_rates <- lacw_scc |>
  dplyr::mutate(recycling_rates = round((`Local authority collected waste - sent for recycling-composting-reuse (tonnes)` / `Total local authority collected waste (tonnes)`* 100), 1)) |>
  dplyr::select(`Financial Year`, recycling_rates)

colnames(recycling_rates) <- c("Financial Year", "% sent for recycling")

recycling_rates |>
  flextable::flextable() |>
  flextable::set_table_properties(layout = "autofit") |>
  flextable::theme_vanilla()
```

The LACW recycling rate is shown in Figure \@ref(fig:districtrecyclingratedistricts) for each of the Suffolk Districts.
These have been subject to some reorganisaition in recent years and so the graphs below show the rates for the combines councils before each merger and the current district.
It is possible that the process of re-organisation has disrupted services to some extent.

From 2019 - 20 Suffolk Coastal District Council and Waveney District Council merge to become East Suffolk Council, and Forest Heath District Council and St Edmundsbury Borough Council merged to become "West Suffolk Council".
Their figures have been aggregated in their respective Councils.

```{r districtrecyclingratedistricts, message = FALSE, fig.alig = "center", fig.height = 8, fig.width = 8, fig.cap = "Source: [ENV18 - Local Authority Collected Waste](https://www.gov.uk/government/statistical-data-sets/env18-local-authority-collected-waste-annual-results-tables)"}
district_plot <- lapply(unique(lacw_suffolk$`Local Authority`), function(authority) {
  district_council <- lacw_suffolk |>
    dplyr::filter(`Local Authority` == authority)
  
  recycling_rates <- district_council |>
    dplyr::mutate(recycling_rates = (`Local authority collected waste - sent for recycling-composting-reuse (tonnes)` / `Total local authority collected waste (tonnes)`) * 100) |>
    dplyr::select(`Financial Year`, recycling_rates)
  
  colnames(recycling_rates) <- c("Financial Year", "% sent for recycling")
  
  p4 <- sccthemes::scc_barchart_pct(
    recycling_rates,
    "Financial Year",
    "% sent for recycling",
    title = authority
  )
  
  p4
})

do.call(gridExtra::grid.arrange, c(district_plot, ncol = 1, top = "Recycling rate by district council"))
```

# Waste Capacity

## Waste Treatment

The capacity of waste management infrastructure has been calculated using the throughput of the facilities in the Plan Area as recorded in the Waste Data Interrogator.
Shown in Table \@ref(tab:wastecapacity) The amount of waste managed in all facilities in Suffolk increased from 2019 to 2020 and in particular the amount of waste managed at the Energy from Waste facility at Great Blakenham increased following the grant of planning permission to enable it to manage 295,000 tonnes per annum, an increase from the original capacity of 269,000 tonnes per annum.

**TZ: In version 4 (as per 05/22) of the 2020 WDI, geographical descriptors have been removed from mobile plant permits. Suffolk totals can therefore not be calculated for 2020. For that reason I have also left 2019 out, to allow for comparison.**

```{r wastecapacity, ft.align = "center", tab.id = "wastecapacity", label = "wastecapacity", tab.cap = "Source: [Waste Data Interrogator 2019](https://find-data-beta.cloudapps.digital/dataset/d409b2ba-796c-4436-82c7-eb1831a9ef25/2019-waste-data-interrogator), [2020](https://www.data.gov.uk/dataset/bb40d091-a346-4b75-aa54-df7d347bed93/2020-waste-data-interrogator) and [2021](https://www.data.gov.uk/dataset/d8a12b93-03ef-4fbf-9a43-1ca7a054479c/2021-waste-data-interrogator)."}
# NOTE: the WDI is published in a .xlsb format, which is notoriously difficult to read.
# A package, {xlsb} does exist with a read_xslb() function, unfortunately at the time of writing not available for my version of R.
# So... the "YEAR Waste Received" sheet was copied into a standalone .xlsx file
# Copy pasting data is generally bad practice (because it is prone to user error),
# But I had no other option in this instance
wdi_totals_table <- wamreport$wdi_totals |>
  tidyr::pivot_wider(names_from = `year`, values_from = `tonnes`)

wdi_totals_table <- wdi_totals_table |>
  dplyr::filter(!`Site Category` %in% c("Landfill", "Storage", "Mobile Plant"))

wdi_totals_table[is.na(wdi_totals_table)] <- 0

wdi_names <- c("Category", "2019", "2020", "2021")
colnames(wdi_totals_table) <- wdi_names

treated_totals <- data.frame(
  `Category` = c("Total waste treated in Suffolk"),
  `2019` = sum(wdi_totals_table$`2019`),
  `2020` = sum(wdi_totals_table$`2020`),
  `2021` = sum(wdi_totals_table$`2021`)
)
colnames(treated_totals) <- wdi_names

totals_formatting <- function(x) {
  x <- format(round(x, 0), big.mark = ",")
  x
}

wdi_totals_table <- dplyr::bind_rows(wdi_totals_table, treated_totals)
wdi_totals_table$`2019` <- totals_formatting(wdi_totals_table$`2019`)
wdi_totals_table$`2020` <- totals_formatting(wdi_totals_table$`2020`)
wdi_totals_table$`2021` <- totals_formatting(wdi_totals_table$`2021`)

wdi_totals_table |>
  flextable::flextable() |>
  flextable::set_table_properties(layout = "autofit") |>
  flextable::add_header_row(
    colwidths = c(4),
    values = c("Waste treated in facilities in Suffolk in 2019, 2020 and 2021 (tonnes)")
  ) |>
  flextable::bold(i = 7, j = 1:4, part = "body") |>
  flextable::theme_vanilla() |>
  flextable::footnote(i = 7, j = 1, 
           value = flextable::as_paragraph(c("Previous reports have included figures for 'Mobile Plants' sites. EA have removed geographical descriptors for this Category, so it can no longer be reported on County level. Therefore totals may be down from previous reports.")),
           ref_symbols = c("*"),
           part = "body")
```

Total waste treated in Table \@ref(tab:wastecapacity) does not include landfill.
Figure \@ref(fig:wastecapacityplot) shows Landfill alongside other categories, to give an idea of scale.

```{r wastecapacityplot, fig.height = 4, fig.width = 8, fig.cap = "Source: [Waste Data Interrogator 2019](https://find-data-beta.cloudapps.digital/dataset/d409b2ba-796c-4436-82c7-eb1831a9ef25/2019-waste-data-interrogator), [2020](https://www.data.gov.uk/dataset/bb40d091-a346-4b75-aa54-df7d347bed93/2020-waste-data-interrogator) and [2021](https://www.data.gov.uk/dataset/d8a12b93-03ef-4fbf-9a43-1ca7a054479c/2021-waste-data-interrogator)."}
wdi_plotted <- wamreport$wdi_totals |>
  dplyr::filter(!`Site Category` %in% c("Mobile Plant", "Storage"))

sccthemes::scc_barchart_grouped(
  wdi_plotted,
  "Site Category",
  "tonnes",
  "year",
  title = "Waste treated in Suffolk (tonnes)"
) +
  ggplot2::scale_fill_manual(values = c("#e2eefa", "#2d6ca2", "#e8850c"))

```

## Waste to Landfill

Figure \@ref(fig:wastetolandfill) shows active Landfill sites.
There is only one active landfill site receiving non-hazardous waste in Suffolk and this is Masons Landfill site which is operated by Valencia Waste Management.
**TZ: This doesn't seem relevant anymore, see Folly Farm**

Shrublands Quarry, an inert landfill site operated by Brett Aggregates also received significant quantities of inert waste in 2019.

Comment on Lawn Farm Quarry

Folly Farm Waste Management 

```{r wastetolandfill, warning = FALSE, fig.alig = "center", fig.height = 6, fig.width = 8, fig.cap = "Source: [Waste Data Interrogator 2019](https://find-data-beta.cloudapps.digital/dataset/d409b2ba-796c-4436-82c7-eb1831a9ef25/2019-waste-data-interrogator), [2020](https://www.data.gov.uk/dataset/bb40d091-a346-4b75-aa54-df7d347bed93/2020-waste-data-interrogator) and [2021](https://www.data.gov.uk/dataset/d8a12b93-03ef-4fbf-9a43-1ca7a054479c/2021-waste-data-interrogator)."}

landfill_totals <- wamreport$totals_by_site
landfill_totals$`Site Name` <- gsub("EPR", "", landfill_totals$`Site Name`)
landfill_totals$`Site Name` <- gsub("BV4517IM", "", landfill_totals$`Site Name`)
landfill_totals$`Site Name` <- gsub("\\/", "", landfill_totals$`Site Name`)
landfill_totals$`Site Name` <- gsub(" - ", "", landfill_totals$`Site Name`)

landfill_totals <- landfill_totals |>
  dplyr::filter(`Site Category` == "Landfill")

landfill_totals <- landfill_totals[order(-landfill_totals$tonnes),]

sccthemes::scc_barchart_grouped(
  landfill_totals,
  "Site Name",
  "tonnes",
  "year",
  title = "Waste Received to Landfill in Suffolk (tonnes)"
) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(size = 10, angle = 45, hjust = 1)) +
  ggplot2::scale_fill_manual(values = c("#e2eefa", "#2d6ca2", "#e8850c"))
```

## Remaining landfill capacity

Permitted landfill operators have a condition in their permits to report the remaining landfill void (capacity) of their sites at the end of the calendar year.
Operators can claim commercial confidentiality.
See Table \@ref(tab:tablc)

```{r tablc, message = FALSE, warning = FALSE, ft.align = "center", tab.id='tablc', label = 'tablc', tab.cap = "Source: [Remaining Landfill Capacity 2019, 2020 and 2021.](https://www.data.gov.uk/dataset/237825cb-dc10-4c53-8446-1bcd35614c12/remaining-landfill-capacity)"}
capacity <- wamreport$rlc_capacity

capacity |>
  flextable::flextable() |>
  flextable::set_table_properties(layout = "autofit") |>
  flextable::add_header_row(
    colwidths = c(6),
    values = c("Remaining landfill capacity reported in Suffolk (cubic metres)")
  ) |>
  flextable::theme_vanilla()
```

In Figure \@ref(fig:landfillcapacityplot) Landfill capacity change is shown between 2019 and 2021, for sites where there is capacity.

```{r landfillcapacityplot, fig.height = 6, fig.width = 8, fig.cap = "Source: [Remaining Landfill Capacity 2019, 2020 and 2021](https://www.data.gov.uk/dataset/237825cb-dc10-4c53-8446-1bcd35614c12/remaining-landfill-capacity)"}
capacity_plot <- tidyr::pivot_longer(
  capacity, 
  cols = c(`2019`, `2020`, `2021`), 
  names_to = "year",
  values_to = "cubic_metres"
)

capacity_plot <- capacity_plot |>
  dplyr::filter(cubic_metres > 0)

sccthemes::scc_barchart_grouped(
  capacity_plot,
  "Facility",
  "cubic_metres",
  "year",
  title = "Remaining Landfill capacity by Facility in Suffolk (cubic metres)") +
  ggplot2::scale_fill_manual(values = c("#e2eefa", "#2d6ca2", "#e8850c")) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(size = 10, angle = 45, hjust = 1))
```

# Waste Movements

```{r waste_managed_in_suffolk, ft.align = "center", tab.id='tab1', label = 'tab1', tab.cap = "Source: [Waste Data Interrogator 2019](https://find-data-beta.cloudapps.digital/dataset/d409b2ba-796c-4436-82c7-eb1831a9ef25/2019-waste-data-interrogator), [2020](https://www.data.gov.uk/dataset/bb40d091-a346-4b75-aa54-df7d347bed93/2020-waste-data-interrogator) and [2021](https://www.data.gov.uk/dataset/d8a12b93-03ef-4fbf-9a43-1ca7a054479c/2021-waste-data-interrogator)."}
wdi_origin <- wamreport$waste_origin

relevant_counties <- c(
  "Suffolk",
  "Norfolk",
  "Essex",
  "Cambridgeshire",
  "Peterborough"
)

counties <- wdi_origin |>
  dplyr::filter(`Origin WPA` %in% relevant_counties)
other <- wdi_origin |>
  dplyr::filter(!`Origin WPA` %in% relevant_counties) |>
  dplyr::select(tonnes, year) |>
  dplyr::group_by(year) |>
  dplyr::summarise(tonnes = sum(tonnes))

other$`Origin WPA` <- "Other"

wdi_origin_table <- dplyr::bind_rows(counties, other)

wdi_origin_table <- wdi_origin_table |>
  tidyr::pivot_wider(names_from = `year`, values_from = `tonnes`)
wdi_origin_table <- wdi_origin_table[(order(-wdi_origin_table$`2020`)),]

suffolk_total <- wdi_origin_table |>
  dplyr::filter(`Origin WPA` == "Suffolk")

from_others <- data.frame(
  `Origin` = c("Total waste managed origin outwith Suffolk"),
  `2019` = sum(wdi_origin_table$`2019`) - suffolk_total$`2019`,
  `2020` = sum(wdi_origin_table$`2020`) - suffolk_total$`2020`,
  `2021` = sum(wdi_origin_table$`2021`) - suffolk_total$`2021`
)

managed_totals <- data.frame(
  `Origin` = c("Total waste managed in Suffolk"),
  `2019` = sum(wdi_origin_table$`2019`),
  `2020` = sum(wdi_origin_table$`2020`),
  `2021` = sum(wdi_origin_table$`2021`)
)
colnames(from_others) <- colnames(wdi_origin_table)
colnames(managed_totals) <- colnames(wdi_origin_table)

wdi_origin_table <- dplyr::bind_rows(wdi_origin_table, from_others)
wdi_origin_table <- dplyr::bind_rows(wdi_origin_table, managed_totals)

wdi_origin_table$`2019` <- totals_formatting(wdi_origin_table$`2019`)
wdi_origin_table$`2020` <- totals_formatting(wdi_origin_table$`2020`)
wdi_origin_table$`2021` <- totals_formatting(wdi_origin_table$`2021`)

wdi_origin_table |>
  flextable::flextable() |>
  flextable::set_table_properties(layout = "autofit") |>
  flextable::add_header_row(
    colwidths = c(4),
    values = c("All waste managed in Suffolk (tonnes)")
  ) |>
  flextable::bold(i = 7:8, j = 1:4, part = "body") |>
  flextable::theme_vanilla()
```

```{r waste_to_landfill2, ft.align = "center", tab.id='tab2', label='tab2', tab.cap = "Source: [Waste Data Interrogator 2019](https://find-data-beta.cloudapps.digital/dataset/d409b2ba-796c-4436-82c7-eb1831a9ef25/2019-waste-data-interrogator), [2020](https://www.data.gov.uk/dataset/bb40d091-a346-4b75-aa54-df7d347bed93/2020-waste-data-interrogator) and [2021](https://www.data.gov.uk/dataset/d8a12b93-03ef-4fbf-9a43-1ca7a054479c/2021-waste-data-interrogator)."}
wdi_landfill <- wamreport$landfill_waste

counties_landfill <- wdi_landfill |>
  dplyr::filter(`Origin WPA` %in% relevant_counties)
other_landfill <- wdi_landfill |>
  dplyr::filter(!`Origin WPA` %in% relevant_counties) |>
  dplyr::select(tonnes, year) |>
  dplyr::group_by(year) |>
  dplyr::summarise(tonnes = sum(tonnes))

other_landfill$`Origin WPA` <- "Other"

wdi_landfill_table <- dplyr::bind_rows(counties_landfill, other_landfill)

wdi_landfill_table <- wdi_landfill_table |>
  tidyr::pivot_wider(names_from = `year`, values_from = `tonnes`)
wdi_landfill_table <- wdi_landfill_table[(order(-wdi_landfill_table$`2020`)),]

wdi_landfill_table[is.na(wdi_landfill_table)] <- 0

landfill_totals <- data.frame(
  `Origin` = c("Landfill total in Suffolk"),
  `2019` = sum(wdi_landfill_table$`2019`),
  `2020` = sum(wdi_landfill_table$`2020`),
  `2021` = sum(wdi_landfill_table$`2021`)
)

colnames(landfill_totals) <- colnames(wdi_landfill_table)

wdi_landfill_table <- dplyr::bind_rows(wdi_landfill_table, landfill_totals)

wdi_landfill_table$`2019` <- totals_formatting(wdi_landfill_table$`2019`)
wdi_landfill_table$`2020` <- totals_formatting(wdi_landfill_table$`2020`)
wdi_landfill_table$`2021` <- totals_formatting(wdi_landfill_table$`2021`)

wdi_landfill_table |>
  flextable::flextable() |>
  flextable::set_table_properties(layout = "autofit") |>
  flextable::add_header_row(
    colwidths = c(4),
    values = c("Waste sent to landfill in Suffolk (tonnes)")
  ) |>
  flextable::bold(i = 6, j = 1:4, part = "body") |>
  flextable::theme_vanilla()

```

The total amount of waste sent **from** Suffolk to other waste planning authority areas is relatively modest as shown in the table below:

```{r waste_from_Suffolk, ft.align = "center", tab.id='tab3', label='tab3', tab.cap = "Source: [Waste Data Interrogator 2019](https://find-data-beta.cloudapps.digital/dataset/d409b2ba-796c-4436-82c7-eb1831a9ef25/2019-waste-data-interrogator), [2020](https://www.data.gov.uk/dataset/bb40d091-a346-4b75-aa54-df7d347bed93/2020-waste-data-interrogator) and [2021](https://www.data.gov.uk/dataset/d8a12b93-03ef-4fbf-9a43-1ca7a054479c/2021-waste-data-interrogator)."}
wdi_export <- wamreport$waste_destination

wdi_export$`Facility WPA`[is.na(wdi_export$`Facility WPA`)] <- "Other"

counties_export <- wdi_export |>
  dplyr::filter(`Facility WPA` %in% relevant_counties)

other_export <- wdi_export |>
  dplyr::filter(!`Facility WPA` %in% relevant_counties) |>
  dplyr::select(tonnes, year) |>
  dplyr::group_by(year) |>
  dplyr::summarise(tonnes = sum(tonnes))

other_export$`Facility WPA` <- "Other"

wdi_export_table <- dplyr::bind_rows(counties_export, other_export)

wdi_export_table <- wdi_export_table |>
  tidyr::pivot_wider(names_from = `year`, values_from = `tonnes`)
wdi_export_table <- wdi_export_table[(order(-wdi_export_table$`2021`)),]

wdi_export_table[is.na(wdi_export_table)] <- 0

export_totals <- data.frame(
  `Origin` = c("Sent from Suffolk total"),
  `2019` = sum(wdi_export_table$`2019`),
  `2020` = sum(wdi_export_table$`2020`),
  `2021` = sum(wdi_export_table$`2021`)
)

colnames(export_totals) <- colnames(wdi_export_table)

wdi_export_table <- dplyr::bind_rows(wdi_export_table, export_totals)

wdi_export_table$`2019` <- totals_formatting(wdi_export_table$`2019`)
wdi_export_table$`2020` <- totals_formatting(wdi_export_table$`2020`)
wdi_export_table$`2021` <- totals_formatting(wdi_export_table$`2021`)

wdi_export_table |>
  flextable::flextable() |>
  flextable::set_table_properties(layout = "autofit") |>
  flextable::add_header_row(
    colwidths = c(4),
    values = c("Waste sent to other authorities from Suffolk (tonnes)")
  ) |>
  flextable::bold(i = 6, j = 1:4, part = "body") |>
  flextable::theme_vanilla()
```

Suffolk largely achieves the goal of net self-sufficiency in the management of the waste arising within its area.
This is demonstrated by the fact that there was a net flow of waste into Suffolk of **`r totals_formatting(from_others$"2019" - export_totals$"2019")` tonnes** in 2019, a net flow into the county of **`r totals_formatting(from_others$"2020" - export_totals$"2020")` tonnes** in 2020 and a net flow out of the county of **`r totals_formatting(export_totals$"2021" - from_others$"2021")` tonnes** This includes non-hazardous and inert wastes.
The net flow is calculated by taking the "Total waste managed origin outwith Suffolk" in Table \@ref(tab:tab1), and subtracting the "Sent from Suffolk total" in Table \@ref(tab:tab3)

## Waste Water Treatment

Of the tonnages sent from Suffolk to Norfolk, a large proportion comprised sludge which was sent to facilities managed by Anglian Water, as highlighted in Table \@ref(tab:tab4).

```{r suffolk_to_norfolk, warning = FALSE, message = FALSE, ft.align = "center", tab.id='tab4', label='tab4', tab.cap = "Source: [Waste Data Interrogator 2019](https://find-data-beta.cloudapps.digital/dataset/d409b2ba-796c-4436-82c7-eb1831a9ef25/2019-waste-data-interrogator), [2020](https://www.data.gov.uk/dataset/bb40d091-a346-4b75-aa54-df7d347bed93/2020-waste-data-interrogator) and [2021](https://www.data.gov.uk/dataset/d8a12b93-03ef-4fbf-9a43-1ca7a054479c/2021-waste-data-interrogator)."}
wdi_norfolk <- wamreport$od_waste

aw2019 <- wdi_norfolk |>
  dplyr::filter(Operator %in% c("Anglian Water Services Limited", "Anglian Water Services Ltd") & year == "2019")

aw2019 <- sum(aw2019$tonnes)

aw2020 <- wdi_norfolk |>
  dplyr::filter(Operator %in% c("Anglian Water Services Limited", "Anglian Water Services Ltd") & year == "2020")

aw2020 <- sum(aw2020$tonnes)

aw2021 <- wdi_norfolk |>
  dplyr::filter(Operator %in% c("Anglian Water Services Limited", "Anglian Water Services Ltd") & year == "2021")

aw2021 <- sum(aw2021$tonnes)

wdi_norfolk$tonnes <- totals_formatting(wdi_norfolk$tonnes)

wdi_norfolk <- tidyr::pivot_wider(wdi_norfolk, names_from = year, values_from = tonnes)
wdi_norfolk <- wdi_norfolk |> 
  dplyr::select(
    `Site Name`,
    Operator,
    `2019`,
    `2020`,
    `2021`
  )

wdi_norfolk[1:7,] |>
  flextable::flextable() |>
  flextable::set_table_properties(layout = "autofit") |>
  flextable::add_header_row(
    colwidths = c(5),
    values = c("Waste sent to Norfolk from Suffolk (tonnes), highlighting Anglian Water")
  ) |>
  #bold(i = c(1,3,6), j = 1:4, part = "body") |>
  flextable::bg(i = c(2,3,7), j = 1:5, part = "body", bg = "#e2eefa") |>
  # color(i = c(1,3,6), j = 1:4, part = "body", color = "white") |>
  flextable::theme_vanilla()
```

In 2019, `r totals_formatting(aw2019)` tonnes of sludges were sent to Whitlingham Sludge Treatment Centre, Thetford Sludge Treatment Centre and King's Lynn Sludge Treatment Centre in Norfolk by Anglian Water.
In 2020, `r totals_formatting(aw2020)` tonnes of sludges were sent to these facilities, and `r totals_formatting(aw2021)` in 2021.

Although in 2020 the total nearly halved, the reduction 2021 has slowed.
This implies that there is still a shortage of waste water treatment capacity in Suffolk and discussions with the Anglian Water would be useful to identify their needs to enable Suffolk to become more self-sufficient in managing this waste stream.

## Thetford Power Station: poultry litter incineration

```{r thetford, warning = FALSE, message = FALSE, ft.align = "center", tab.id='tab5', label='tab5', tab.cap = "Source: [Waste Data Interrogator 2019](https://find-data-beta.cloudapps.digital/dataset/d409b2ba-796c-4436-82c7-eb1831a9ef25/2019-waste-data-interrogator), [2020](https://www.data.gov.uk/dataset/bb40d091-a346-4b75-aa54-df7d347bed93/2020-waste-data-interrogator) and [2021](https://www.data.gov.uk/dataset/d8a12b93-03ef-4fbf-9a43-1ca7a054479c/2021-waste-data-interrogator)."}

wdi_thetford <- wamreport$wdi_all |> 
  dplyr::filter(`Origin WPA` == "Suffolk" & `Facility WPA` == "Norfolk" & Operator == "EPR Thetford Limited")

wdi_norfolk[1:3,] |>
  flextable::flextable() |>
  flextable::set_table_properties(layout = "autofit") |>
  flextable::add_header_row(
    colwidths = c(5),
    values = c("Waste sent to Norfolk from Suffolk (tonnes), highlighting Thetford Power Station")
  ) |>
  #bold(i = c(1,3,6), j = 1:4, part = "body") |>
  flextable::bg(i = c(1), j = 1:5, part = "body", bg = "#e2eefa") |>
  # color(i = c(1,3,6), j = 1:4, part = "body", color = "white") |>
  flextable::theme_vanilla()
```

One remarkable rise from 2020 to 2021, is the increase in tonnage sent to Thetford Power Station, operated by EPR Thetford Limited.
This waste falls under SOC Category "09 - Animal and vegetable wastes" and SOC Sub Category "Slurry and manure", the waste is described as "animal faeces, urine and manure (including spoiled straw), effluent, collected separately and treated off-site".
Thetford Power Station is an Animal By-Products Incinerator, [which converts poultry litter into electricity](https://www.mreuk.com/thetford).

# Performance of the Waste Policies and Conclusions

The outcome of the waste planning process in Suffolk has been to deliver broadly sufficient capacity of waste management facilities to manage the waste arising within the County.This includes a wide variety of treatment facilities to manage different types of materials, an energy from waste facility to manage residual waste from household, commercial and industrial sources, and sufficient landfill to dispose of the remainder.

Efforts to reduce, reuse and recycle waste will continue, particularly in the context of the implementation of the Government's Resources and Waste Strategy which was published in 2018.
This proposes more separate collection of food waste and the separate collection of other easily recyclable materials and so consideration will need to be given to the delivery of bulking and transfer stations that may be associated with this.

Consideration may need to be given to the development of additional waste water treatment capacity, given the significant movements of sludges out of the Plan Area and liaison with Anglian Water will be needed to take this forward.

\newpage

# Appendix
